Deep Embedded Clustering for Open Cluster Characterization with Gaia DR2 Data

### 2021-03-06 paper/paper.tex
paper: Update title
paper: Improve results section
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

\documentclass[11pt,a4paper,english,twocolumn]{article}
\usepackage{unir_paper}
\usepackage{subcaption}
\usepackage{tablefootnote}

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
\author{CD. Álvaro, C. Guzman, J. Álvaro}
\date{24th of January, 2021}

@@

\subsection{Selection of features}
\label{sec:feature_selection}

@@

@@

@@

%% Comparación con K-Means, DEC y DEC filter
\begin{figure*}[!htbp]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_pm_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_hr_diagram_melotte_25.png}
      \caption{K-Means}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
      \caption{DEC}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_filtered_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_filtered_melotte_25.png}
      \caption{DEC (filt.)}
  \end{subfigure}
  \caption{Evolution of Melotte 25 characterization}
  \label{fig:melotte_25_characterization_evolution}
\end{figure*}

We have tested our model with the Melotte 25 dataset, which is one of the most difficult dataset
to characterize OC by its distribution and extension, because it is located in a region with
400,000 stars.

In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, we compared
our model against a K-Means classification algorithm. In addition, we perform a second execution
only in our model where we filtered the quantiles in the Melotte 25 dataset.

Our model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter
Notebooks~\cite{Kluyver2016jupyter}. All tests were run on an Apple Mac Pro Late 2013 with a
2.7GHz 12-Core Intel Xeon E5-2697v2, and 64GB RAM 1866MHz DDR3. For the GPU, we use a graphic
card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available
at \href{https://github.com/cdalvaro/machine-learning-master-thesis}
{https://github.com/cdalvaro/machine-learning-master-thesis}}.

% More results are available at Appendix \ref{sec:appendix}.

Figure~\ref{fig:melotte_25_characterization_evolution} shows the evolution of the characterization
of Melotte 25. The first column corresponds to the characterization obtained by applying K-Means
(with N=10) alone to the dataset. The middle column is the characterization made by our model.
And the column on the right shows the characterization made by our model but filtering the
quantiles lower than 0.10 and higher than 0.90. The top row shows the proper motion configuration
space, while the bottom one shows the Hertzsprung–Russell diagrams.

@@

The clustering made by K-Means arises 9 groups, but most of them contain stars that do not belong
to the Melotte 25 OC. However, after applying our model, most of the stars are grouped into a single
cluster. One of the other groups is the OC we are looking for, and the other seven could be new OCs
inside the studied region.

The dataset Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in
declination with a radius of 165 arcmin. As it was mentioned before, this region contains more than
400,000 stars making it difficult to group stars of the same region.

@@

By comparing the results obtained with our model, we can see that they are in range with those extracted
from SIMBAD astronomical database~\cite{wenger2000simbad}. While the results provided by K-Means model are
far from the official data. This comparison demonstrates that our model improves the results obtained with
a raw K-Means.

\begin{table}[!htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{l|c|c|c|c}
        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
        \hline
        \textbf{Simbad}\tablefootnote{Results have been taken from the SIMBAD astronomical database~\cite{wenger2000simbad}}
        & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
        %Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
        K-Means & 79.936 \( \pm \) 3.7 & -45.362 \( \pm \) 3.8 & 20.901 \( \pm \) 0.31 & 374 \\
        DECOCC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
        \textbf{DECOCC (filt.)} & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
      \end{tabular}
    }
    \caption{Summary for the Melotte 25 results.}
    \label{tab:results_melotte_25}
  \end{center}
\end{table}

Figure~\ref{fig:result_melotte_25_dec} shows the groups found using our model.
Our model labels Melotte 25 as group \emph{g5}.

\begin{figure*}[!hbt]
  \centering
  \begin{subfigure}[t]{0.30\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.30\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_25/dec_parallax_melotte_25.png}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.30\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
  \end{subfigure}
  \caption{Melotte 25 characterization.
           From left to right: proper motion configuration space,
           parallax histogram and H-R diagram.}
  \label{fig:result_melotte_25_dec}
\end{figure*}

Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters
used for the characterization.

\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c}
      \textbf{Hyperparameter} & \textbf{Value} \\
      \hline
      Number of Clusters & 9 \\
      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
      Kernel Initializer Seed & 11 \\
      Quantil Threshold & 0.1 \\
    \end{tabular}
    \caption{Hyperparameters for Melotte 25.}
    \label{tab:hyperparameters_melotte_25}
  \end{center}
\end{table}

These hyperparameters are used to fine-tuning our model and improve its accuracy. Although they depend on
the studied region, no one of them is related with physical properties of the cluster. That makes easier
to automate the selection of hyperparameters.

\paragraph{Final Discussion.}

We have applied our proposed method to clusters with different typologies:

\begin{itemize}
  \item NGC 2516 (Ap.~\ref{sec:ngc2516}) has it proper motion center deviated from the origin but
        it is embedded inside a big cloud of stars with similar proper motions.
  \item NGC 2632 (Ap.~\ref{sec:ngc2632}), in addition to Melotte 22, is a cluster whose proper motion center
        is not located at the origin and has a well separated parallax center.
  \item NGC 2682 (Ap.~\ref{sec:ngc2682}), on the other hand, has its parallax centered within the region's gaussian,
        which complicates its detection although its proper motion center is deviated from the origin.
  \item Melotte 22 (Ap.~\ref{sec:melotte22}), as well as Melotte 25, is a cluster closer to us than the other.
        That makes its membership stars to be more scattered than previous clusters which are more compact.
\end{itemize}

All these clusters have a wide variety of diameters, from 25 to 330 arcmin,
as well as the number of stars that belong to them. NGC 2682 has 3,000 stars
while Melotte 25 is located inside a region with more than 400,000 stars.
More information about the studied clusters is available in Table~\ref{tab:clusters_summary}.

\begin{table}[htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{l|c|c|c|c}
        \textbf{OC} & \emph{\( \alpha \) J2000 \((degrees)\)} & \emph{\( \delta \) J2000 \((degrees)\)}
        & \emph{Radius \((arcmin)\)} & \emph{\# stars} \\
        \hline
        NGC 2516 & 119.517 & -60.753 & 15 & 12,869 \\
        NGC 2632 & 130.1 & 19.667 & 35 & 13,167 \\
        NGC 2682 & 132.825 & 11.8 & 12.5 & 2,839 \\
        Melotte 22 & 56.75 & 24.117 & 60 & 61,552 \\
        Melotte 25 & 66.725 & 15.867 & 165 & 433,996 \\
      \end{tabular}
    }
    \caption{Right ascension, declination, radius and number of stars of studied clusters.
             The number of stars corresponds to those stars contained within a cone of center
             \((\alpha, \delta)\) and radius the cluster's radius multiplied by a factor of 1.5.}
    \label{tab:clusters_summary}
  \end{center}
\end{table}

In all cases, the model has resolved properly the identity of the cluster
and has characterized the membership stars, giving compatible results
with the ones obtained with classic procedures and VO tools.

@@

Only at the end of the characterization, a fine tuning driven by the user,
can be applied in order to improve the selection by discarding those
stars which fall outside the given quantiles.

@@

\appendix
\section{Appendix}
\label{sec:appendix}

\subsection{NGC 2516}
\label{sec:ngc2516}

\begin{table}[htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{l|c|c|c|c}
        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
        \hline
        \textbf{Simbad} & -4.6579 \( \pm \) 0.0075 & 11.1517 \( \pm \) 0.0075 & 2.4118 \( \pm \) 0.0006 & 1727 \\
        Clusterix & -4.652 \( \pm \) 0.523 & 11.203 \( \pm \) 0.454 & 2.409 \( \pm \) 0.127 & 638 \\
        K-Means & -4.344 \( \pm \) 0.14 & 9.507 \( \pm \) 0.19 & 2.268 \( \pm \) 0.01 & 1542 \\
        DEC & -4.426 \( \pm \) 0.17 & 9.952 \( \pm \) 0.20 & 2.436 \( \pm \) 0.01 & 1532 \\
        \textbf{DEC (filt.)} & -4.502 \( \pm \) 0.14 & 10.114 \( \pm \) 0.17 & 2.392 \( \pm \) 0.004 & 1072 \\
      \end{tabular}
    }
    \caption{NGC 2516 results.}
    \label{tab:app_results_ngc_2516}
  \end{center}
\end{table}

Table~\ref{tab:app_results_ngc_2516} shows a results summary for NGC 2516.
The top row of Figure~\ref{fig:app_result_ngc_2516_clusterix_kmeans} shows
the characterization using Clusterix+TOPCAT tools while the bottom row
shows eight clusters found by K-Means.

>>>>

\documentclass[11pt,a4paper,english,twocolumn]{article}
\usepackage{unir_paper}
\usepackage{subcaption}
\usepackage{tablefootnote, multirow}

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Deep Embedded Clustering for Open Cluster Characterization with Gaia DR2 Data}
\author{CD. Álvaro, C. Guzman, J. Álvaro}
\date{24th of January, 2021}

@@

\subsection{Selection of Features}
\label{sec:feature_selection}

@@

%% Comparación con K-Means, DEC y DEC filter
\begin{figure*}[!htp]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_pm_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_hr_diagram_melotte_25.png}
      \caption{K-Means}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
      \caption{DECOCC}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_filtered_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_filtered_melotte_25.png}
      \caption{DECOCC (filt.)}
  \end{subfigure}
  \caption{Evolution of Melotte 25 characterization}
  \label{fig:melotte_25_characterization_evolution}
\end{figure*}

@@

\newpage

@@

We performed two tested of our model:

\begin{enumerate}
  \item with the Melotte 25 dataset, which is one of the most difficult dataset to characterize OC
        by its distribution and extension, because it is located in a region with 400,000 stars.
  \item with a variety of clusters of different typologies.
\end{enumerate}

In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, in the both
tests we compared our model against a K-Means classification algorithm. In addition, we perform a
second execution only in our model where we filtered the quantiles in the Melotte 25 dataset.

Our DECOCC model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter
Notebooks~\cite{Kluyver2016jupyter}. All tests were run on an Apple Mac Pro Late 2013 with a
2.7GHz 12-Core Intel Xeon E5-2697v2, and 64GB RAM 1866MHz DDR3. For the GPU, we use a graphic
card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available
at \href{https://github.com/cdalvaro/machine-learning-master-thesis}
{https://github.com/cdalvaro/machine-learning-master-thesis}}.

\subsection{Melotte 25 dataset}

Figure~\ref{fig:melotte_25_characterization_evolution} shows the evolution of the characterization
of Melotte 25. The first column corresponds to the characterization obtained by applying only
K-Means (with N=10). The middle column shows the result of our model. And the last column describes
the characterization made by our model but filtering the quantiles lower than 0.10 and higher than
0.90. The top row shows the proper motion configuration space, while the bottom one shows the
Hertzsprung–Russell diagrams.

@@

The clustering made by K-Means arises 9 groups, but most of them contain stars that do not belong
to the Melotte 25 OC. However, in our model, the stars regroup and although we have the same 9
groups of K-Means, almost all of the stars are assigned to a group that would contain the stars
that would not belong to the OC. The remaining groups contain either the OC we are looking for or
other possible new OCs inside the studied region. Moreover, if we exclude those stars that remain
below 0.10 and above 0.90 percentiles, and execute again this refinement in our model (third column
in Figure~\ref{fig:melotte_25_characterization_evolution}), we observe an improvement in the
results of our DECOCC model.

\begin{figure}[!hbt]
  % \centering
  % \begin{subfigure}[t]{0.30\textwidth}
  %   \centering
  %   \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[t]{0.30\textwidth}
    \centering
    \includegraphics[width=0.36\textwidth]{../figures/melotte_25/dec_parallax_melotte_25.png}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[t]{0.30\textwidth}
  %   \centering
  %   \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
  % \end{subfigure}
  % \caption{Melotte 25 characterization.
  %          From left to right: proper motion configuration space,
  %          parallax histogram and H-R diagram.}
  \caption{Parallax histogram of Melotte 25 characterization.}
  \label{fig:result_melotte_25_dec}
\end{figure}

The dataset Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in
declination with a radius of 165 arcmin. As it was mentioned before, this region contains more than
400,000 stars making it difficult to group stars of the same region. Our model labels Melotte 25 as
group \emph{g5}. Figure~\ref{fig:result_melotte_25_dec} shows the parallax histogram.

\begin{table}[!htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{l|c|c|c|c}
        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
        \hline
        Simbad & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
        % Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
        K-Means & 79.936 \( \pm \) 3.7 & -45.362 \( \pm \) 3.8 & 20.901 \( \pm \) 0.31 & 374 \\
        DECOCC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
        DECOCC (filt.) & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
      \end{tabular}
    }
    \caption{Summary for the Melotte 25 results.}
    \label{tab:results_melotte_25}
  \end{center}
\end{table}

@@

By comparing the results obtained with our DECOCC model, we can see that they are in range with those
extracted from Simbad astronomical database~\cite{wenger2000simbad}. While the results provided by K-Means
model are far from the official data. This comparison demonstrates that our model improves the results
obtained with a raw K-Means.

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{l|c}
      \textbf{Hyperparameter} & \textbf{Value} \\
      \hline
      Number of Clusters & 9 \\
      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
      Kernel Initializer Seed & 11 \\
      Quantil Threshold & 0.1 \\
    \end{tabular}
    \caption{Hyperparameters for Melotte 25.}
    \label{tab:hyperparameters_melotte_25}
  \end{center}
\end{table}

Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for the characterization. We used
them to fine-tuning our model and improve its accuracy. Although they depend on the studied region, none of
them are related with physical properties of the cluster, which makes it easy to automate the selection of
hyperparameters.

\subsection{Clusters with different typologies}

We perform now more exhaustive tests of our DECOCC model using open clusters with different typologies (e.g.,
large, small, more extensive, or more concentrated within the region to be analyzed):

\begin{itemize}
  \item NGC 2516, which has the proper motion center deviated from the origin but embedded within a large cloud
        of stars with similar proper motions.
  \item NGC 2632, which, in addition to Melotte 22, is a cluster whose proper motion center is not at the origin
        and has a well-separated parallax center.
  \item NGC 2682, which has the parallax centered within the region's Gaussian, making it difficult its detection.
        Moreover, same as NGC 2516, the proper motion center is deviated from the origin.
  \item Melotte 22, as well as Melotte 25, is a cluster closer to our galaxy. That causes its membership stars to
        be more scattered than previous clusters which are more compact.
\end{itemize}

Table~\ref{tab:clusters_summary} describes more specific information about the previous clusters (we also include
information about the Melotte 25 open cluster, which was previously analyzed). The first column shows the name of
the cluster. The remaining columns are the right ascension, declination, radius, and the number of stars of studied
clusters. The column \# stars correspond to those stars contained within a cone of center \((\alpha, \delta)\) and
radius the cluster's radius multiplied by a factor of 1.5.

\begin{table}[htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{l|c|c|c|c}
        \textbf{OC} & \emph{\( \alpha \) J2000 \((degrees)\)} & \emph{\( \delta \) J2000 \((degrees)\)}
        & \emph{Radius \((arcmin)\)} & \emph{\# stars} \\
        \hline
        NGC 2516 & 119.517 & -60.753 & 15 & 12,869 \\
        NGC 2632 & 130.1 & 19.667 & 35 & 13,167 \\
        NGC 2682 & 132.825 & 11.8 & 12.5 & 2,839 \\
        Melotte 22 & 56.75 & 24.117 & 60 & 61,552 \\
        Melotte 25 & 66.725 & 15.867 & 165 & 433,996 \\
      \end{tabular}
    }
    \caption{Properties of clusters for further testing of our model.}
    \label{tab:clusters_summary}
  \end{center}
\end{table}

%% Tabla comparativa de todos los open clusters
\begin{table*}[!htb]
  \begin{center}
    % \resizebox{\columnwidth}{!}{
      \begin{tabular}{c|l|c|c|c|c}
         & Method & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
        \hline
        \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2516}}
        & Simbad & -4.6579 \( \pm \) 0.0075 & 11.1517 \( \pm \) 0.0075 & 2.4118 \( \pm \) 0.0006 & 1727 \\
        & K-Means & -4.344 \( \pm \) 0.14 & 9.507 \( \pm \) 0.19 & 2.268 \( \pm \) 0.01 & 1542 \\
        & DECOCC & -4.426 \( \pm \) 0.17 & 9.952 \( \pm \) 0.20 & 2.436 \( \pm \) 0.01 & 1532 \\
        & DECOCC (filt.) & -4.502 \( \pm \) 0.14 & 10.114 \( \pm \) 0.17 & 2.392 \( \pm \) 0.004 & 1072 \\\\

        % \caption{NGC 2632 results.}
        \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2632}} & Simbad & -36.047 \( \pm \) 0.110 & -12.917 \( \pm \) 0.066 & 5.371 \( \pm \) 0.003 & - \\
        & K-Means & -26.352 \( \pm \) 0.82 & -15.828 \( \pm \) 0.76 & 5.394 \( \pm \) 0.03 & 629 \\
        & DECOCC & -20.012 \( \pm \) 0.69 & -14.742 \( \pm \) 0.58 & 4.686 \( \pm \) 0.03 & 894 \\
        & DECOCC (filt.) & -21.571 \( \pm \) 0.74 & -14.234 \( \pm \) 0.61 & 4.719 \( \pm \) 0.03 & 714 \\\\

        % \caption{NGC 2682 results.}
         \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2682}} & Simbad & -10.9737 \( \pm \) 0.0064 & -2.9396 \( \pm \) 0.0063 & 1.1325 \( \pm \) 0.0011 & 1194 \\
        & K-Means & -8.616 \( \pm \) 0.15 & -3.710 \( \pm \) 0.16 & 1.196 \( \pm \) 0.01 & 1374 \\
        & DECOCC & -8.926 \( \pm \) 0.15 & -3.550 \( \pm \) 0.15 & 1.144 \( \pm \) 0.005 & 1238 \\
        & DECOCC (filt.) & -9.619 \( \pm \) 0.13 & -3.317 \( \pm \) 0.13 & 1.140 \( \pm \) 0.003 & 990 \\\\

        % \caption{Melotte 22 results.}
        \multirow{4}{*}{\rotatebox[origin=c]{90}{Melotte 22}} & Simbad & 19.997 \( \pm \) 0.127 & -45.548 \( \pm \) 0.101 & 7.364 \( \pm \) 0.005 & 1326 \\
        & K-Means & 20.25 \( \pm \) 0.95 & -38.01 \( \pm \) 1.08 & 7.23 \( \pm \) 0.06 & 1378 \\
        & DECOCC & 23.67 \( \pm \) 1.29 & -46.23 \( \pm \) 1.50 & 8.04 \( \pm \) 0.09 & 878 \\
        & DECOCC (filt.) & 19.50 \( \pm \) 0.41 & -44.23 \( \pm \) 0.39 & 7.42 \( \pm \) 0.005 & 438 \\

      \end{tabular}
    % }
    \caption{Summary of the results for the clusters with different typologies.}
    \label{tab:app_results_ngc_2516}
  \end{center}
\end{table*}

As we can see, all those clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number
of stars that belong to them. For instance, NGC 2682 has 3,000 stars while Melotte 25 is inside a region with more
than 400,000 stars.

Table~\ref{tab:app_results_ngc_2516} describes a summary of the results for the clusters with different typologies.
Again, our DECOCC model presents similar results from those extracted from the astronomical database Simbad. Except
for the open cluster NGC 2632. In this cluster, the K-Means presents better results than our model. And this is so
because the center of the proper motion is not at the origin and generates a more separate center of parallax, which
makes it difficult to characterize the open cluster in our model. However, our model presents more robust results in
more difficult open clusters, such as the NGC 2682 that has the parallax centered within the region's Gaussian.

%% comparativa de las figuras de los open clusters
\begin{figure*}[!hbt]
  \rotatebox[origin=c]{90}{{\bfseries NGC 2516}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2516} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2632}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2632} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
    \caption{K-Means}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
    \caption{DECOCC}
  \end{subfigure}%
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
    \caption{DECOCC (filt.)}
  \end{subfigure}
  \caption{Summary of the characterization for the clusters NGC 2516 and NGC 2632.}
  \label{fig:ngc_2516_ngc_2632_characterization_evolution}
\end{figure*}

\begin{figure*}[!hbt]
  \rotatebox[origin=c]{90}{{\bfseries NGC 2682}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_parallax_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_filtered_ngc_2682.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2682} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_hr_diagram_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_filtered_ngc_2682.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries Melotte 22}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries Melotte 22} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
    \caption{K-Means}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
    \caption{DECOCC}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
    \caption{DECOCC (filt.)}
  \end{subfigure}
  \caption{Summary of the characterization for the clusters NGC 2682 and Melotte 22.}
  \label{fig:melotte_22_ngc_2682_characterization_evolution}
\end{figure*}

Figures~\ref{fig:ngc_2516_ngc_2632_characterization_evolution} and~\ref{fig:melotte_22_ngc_2682_characterization_evolution}
show the groups found for the three models. For each open cluster, we describe the parallax and hr diagram.

K-Means and DECOCC have labeled NGC 2516 as \emph{g3}. Although in general, groups between K-Means and our
model may not match.

%% tabla comparativa de los hyperparametros
\begin{table}[htbp]
  \begin{center}

    % \resizebox{\columnwidth}{!}{
        \begin{tabular}{l|l|c}
          & \textbf{Hyperparameter} & \textbf{Value} \\
          \hline
          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2516}} & Number of Clusters & 8 \\
          & Clustering Layer & \(\left[ 50, 50, 60 \right]\) \\
          & Kernel Initializer Seed & 2 \\
          & Quantil Threshold & 0.15 \\\\

          % \caption{NGC 2632 DEC hyperparameters.}
          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2632}} & Number of Clusters & 10 \\
          & Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
          & Kernel Initializer Seed & 10 \\
          & Quantil Threshold & 0.1 \\\\

         % \caption{NGC 2682 DEC hyperparameters.}
          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2682}} & Number of Clusters & 10 \\
          & Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
          & Kernel Initializer Seed & 0 \\
          & Quantil Threshold & 0.1 \\\\

          % \caption{Melotte 22 DEC hyperparameters.}
          \multirow{4}{*}{\rotatebox[origin=c]{90}{Melotte 22}} & Number of Clusters & 5 \\
          & Clustering Layer & \(\left[ 50, 50, 200 \right]\) \\
          & Kernel Initializer Seed & 11 \\
          & Quantil Threshold & 0.1 \\

        \end{tabular}
    % }
    \caption{Hyperparameters used by DECOCC model for the open clusters with different typologies.}
    \label{tab:app_hyperparameters_ngc_2516}
  \end{center}
\end{table}

Table~\ref{tab:app_hyperparameters_ngc_2516} shows the hyperparameters used for characterizing the
open clusters with our method.

In all cases, the DECOCC model has resolved properly the identity of the cluster and has characterized
the membership stars, giving compatible results with the ones obtained with classic procedures and VO
tools.

@@

Only at the end of the characterization, a fine-tuning driven by the user, can be applied in order to
improve the selection by discarding those stars which fall outside the given quantiles.

@@
